---
title: "Exercise Answers"
author: "John Blischak and Emily Davenport"
date: "September 13, 2016"
output: html_document
---

# Lesson 01: 

## Slicing (Subsetting) Data

### Question
A subsection of a data frame is called a [slice]({{ site.github.url }}/reference/#slice).
We can take slices of character vectors as well:

```{r}
animal <- c("m", "o", "n", "k", "e", "y")
# first three characters
animal[1:3]
# last three characters
animal[4:6]
```

1.  If the first four characters are selected using the slice `animal[1:4]`, how can we obtain the first four characters in reverse order?

2.  What is `animal[-1]`?
    What is `animal[-4]`?
    Given those answers,
    explain what `animal[-1:-4]` does.

3.  Use a slice of `animal` to create a new character vector that spells the word "eon", i.e. `c("e", "o", "n")`.

### Solution

1.

```{r}
animal[4:1]
```

2. 
```{r}
animal[-1]
```

```{r}
animal[-4]
```

```{r}
animal[-1:-4]
```

-1:-4 creates a slice from -1 to -4: -1, -2, -3, -4.
Those numbers are then indexed after animal: animal[c(-1, -2, -3, -4)], which results in the first four entries of animal being removed. 

## Plotting Data

### Question

Create a plot showing the standard deviation of the inflammation data for each day across all patients.

### Solution

```{r, eval=FALSE}
sd_day_inflammation <- apply(dat, 2, sd)
plot(sd_day_inflammation)
```


# Lesson 02: Functions

## Functions to Create Graphs

### Question

Write a function called `analyze` that takes a filename as a argument
and displays the three graphs produced in the [previous lesson][01] (average, min and max inflammation over time).
`analyze("data/inflammation-01.csv")` should produce the graphs already shown,
while `analyze("data/inflammation-02.csv")` should produce corresponding graphs for the second data set.
Be sure to document your function with comments.

### Solution
```{r}
analyze <- function(filename) {
# Plots the average, min, and max inflammation over time.
# Input is character string of a csv file.
  dat <- read.csv(file = filename, header = FALSE)
  avg_day_inflammation <- apply(dat, 2, mean)
  plot(avg_day_inflammation)
  max_day_inflammation <- apply(dat, 2, max)
  plot(max_day_inflammation)
  min_day_inflammation <- apply(dat, 2, min)
  plot(min_day_inflammation)
}

```


